---
title: "RWorksheet_Freires#2"
output: pdf_document
date: "2024-09-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create a vector using : operator
a) Sequence from -5 to 5. Write the R code and its output.
Describe its output.

```{r}
num <- -5:5
print(num)
```
- The output would be sequenced by -5 until 5.

b) What will be the value of x?
  
```{r}
x <- 1:7
print(x)
```
2. Create a vector using seq() function
a. seq(1, 3, by=0.2) # specify step size
Write the R script and its output. 

```{r}
num <- seq(1,3, by = 0.2)
print(num)
```
Describe the output.

- The output is a sequenced vector that starts wtih 1 until 3. 

3. A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages:
```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
```

a. Access 3rd element, what is the value?
```{r}
third <- ages[3]
print(third)
```
b. Access 2nd and 4th element, what are the values?
```{r}
sec_fourth <- ages[c(2,4)]
print(sec_fourth)
```
c. Access all but the 4th and 12th element is not
included. Write the R script and its output.

```{r}
all <- ages[-c(4,12)]
print(all)
```
4. Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the
vector, names(x).

```{r}
x <- c("first" = 3, "second" = 0, "third" = 9)
print(x)
```
a. Print the results. Then access x[c("first", "third")].

```{r}
names <- x[c("first", "third")]
print(names)
```
Describe the output.

- The output is showing 3 vectors and then accessing elements of the first and third by giving names. 

5. Create a sequence x from -3:2.
a. Modify 2nd element and change it to 0;
x[2] <- 0
x

```{r}
x <- -3:2
x[2] <- 0
print(x)
```
Describe the output.

- The output shows the sequence from -3 until 2, After accessing the second element, the -2 has changed to 0. 

6. The following data shows the diesel fuel purchased by Mr. Cruz.

Month Jan Feb March Apr May June
Price per liter (PhP) 52.50 57.25 60.00 65.00 74.25 54.00
Purchase–quantity(Liters) 25 30 40 50 10 45

a. Create a data frame for month, price per liter (php)
and purchase-quantity (liter). Write the R scripts and
its output.

```{r}
month <- c("Jan", "Feb", "March", "Apr", "May", "June")
php <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
liter <- c(25, 30, 40, 50, 10, 45)

fuel_data <- data.frame(Month = month, Php = php, Liter = liter)
print(fuel_data)
```

b. What is the average fuel expenditure of Mr. Cruz from
Jan to June? Note: Use ‘weighted.mean(liter,
purchase)‘. Write the R scripts and its output.

```{r}
average_exp <- weighted.mean(php, liter)
print(average_exp)
```
7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths
(in miles) of 141 “major” rivers in North America, as compiled by the US Geological
Survey”.

a. Type “rivers” in your R console.
Create a vector data with 7
elements, containing the number of elements (length)
in rivers, their sum (sum), mean (mean),
median(median), variance(var), standard deviation(sd),
minimum (min) and maximum (max).

```{r}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
print(data)
```

b. What are the results?

- [1]    141.0000 (length of rivers)  83357.0000(sum of rivers)    591.1844(mean of rivers)   425.0000(median of rivers) 243908.4086 (variance of rivers)  493.8708(standard deviation of rivers)    135.0000(minimum length of rivers)   3710.0000(maximum length of rivers)

8. The table below gives the 25 most powerful celebrities and their annual pay as ranked by the editions of Forbes magazine and as listed on the Forbes.com website.


a. Create vectors according to the above table.
Write the R scripts and its output.

```{r}
celebrity <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", " Bruce Springsteen", "Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", "David Letterman", "Phil Mickelson", "J.K Rowling", "Brad Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant")
pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)

celebrity_data <- data.frame(Celebrity = celebrity, Pay = pay)
print(celebrity_data)
```
b. Modify the power ranking and pay of J.K. Rowling.
Change power ranking to 15 and pay to 90. Write the
R scripts and its output.

```{r}

celebrity_data[celebrity_data$celebrity == "J.K. Rowling", "Pay"] <- 90
celebrity_data$Power_ranking <- 1:nrow(celebrity_data)
celebrity_data[celebrity_data$celebrity == "J.K. Rowling", "Power_ranking"] <- 15
print(celebrity_data)

```
c. Create an excel file from the table above and save it
as csv file(PowerRanking). Import the csv file into
the RStudio. What is the R script?

```{r}
write.csv(celebrity_data, "power_ranking.csv", row.names = FALSE)
import_data <- read.csv("power_ranking.csv")
print(import_data)
```

d. Access the rows 10 to 20 and save it as Ranks.RData.
Write the R script and its output.

```{r}
ranks <- celebrity_data[10:20, ]
save(ranks, file = "Ranks.RData")
print(ranks)

```
e. Describe its output.

- The output shows the ranking and pay of top celebrities. 