---
title: "RWorksheet_Freires#1"
output: pdf_document
date: "2024-09-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. a. How many data points?
- 34 data points

b. Write the R code and its output
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41.)
points <- length(age)
print(points)

```
2. Find the reciprocal of the values for age.
```{r}
rec_age <-1 /age
print(rec_age)
```
3. What happen to the new_age?
```{r}
new_age <- c(age, 0, age)
print(new_age)
```
4. Sort the values for age.
```{r}
age <- sort(age)
print(age)
```
5. Find the minimum and maximum value for age.
```{r}
minimum_age <- min(age)
maximum_age <- max(age)
print(minimum_age)
print(maximum_age)
```
6. a. How many data points?
- 12 data points

b. Write the R code and its output
```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
points <- length(data)
print(points)

```
7. Generates a new vector for data where you double every value of data. 

```{r}
double_d <- data * 2
print(double_d)
```
What happen to the data?
- The data increased by double


8. Generate a sequence for the following scenario:
8.1 Integers from 1 to 100.
```{r}
integer <- seq(data)
1:100
```
8.2 Numbers from 20 to 60
```{r}
numbers <- seq(data)
20:60
```

8.3 Mean of numbers from 20 to 60
```{r}
mean_num <- mean(20:60)
print (mean_num)
```


8.4 Sum of numbers from 51 to 91
```{r}
num <- sum(51:91)
total <- sum(num)
print(total)
```

8.5 Integers from 1 to 1,000
```{r}
seq <- 1:1000
print(seq)
```

a. How many data points from 8.1 to 8.4?
```{r}
length1 <- length(seq)
length2 <- length(numbers)
length3 <- length(mean_num)
length4 <- length(num)

sum(length1+length2+length3+length4)

```
b. Write the R code and its output from 8.1 to 8.4
```{r}
seq <- seq(1:100)
numbers <- seq(20:60)
mean_num <-(20:60)
num <-sum(51:91)
```
c. For 8.5 find only maximum data points util 10
```{r}
max <- 1:1000
answer <- max(max[t <- 10])
answer
```

9. Print a vector with the integers between 1 and 100 that are not divisible by 3, 5 and 7 using filter option.
```{r}
Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(100))
```
10. Generate a sequence backwards of the integers from 1 to 100.
```{r}
backward <- seq(100, 1, by = -1)

print(backward)
```
11. List all the natural numbers below 25 that are multiples of 3 or 5. Find the sum of these multiples.
```{r}
numbers <- 1:24
multiples <- numbers[numbers %% 3 == 0 | numbers %% 5 == 0]
print(multiples)

sum <- sum(multiples)
print(sum)

```

12. Statements can be grouped together using braces ‘{’ and ‘}’. A group of statements
is sometimes called a block. Single statements are evaluated when a new line is typed
at the end of the syntactically complete statement. Blocks are not evaluated until a
new line is entered after the closing brace.

x <- {0 + x + 5 }

Describe the output
- The output is an error.


13. Find x[2] and x[3].
```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75,
75, 77)
x2 <- score[2]
x3 <- score[3]
print(x2)
print(x3)
```
14. Create a vector a = c(1,2,NA,4,NA,6,7).
a. Change the NA to 999 using the codes print(a,na.print="-999").
```{r}

a <- c(1,2,NA,4,NA,6,7)

print(a, na.print = "-999")
```
Describe the output

+ the vector contains character that replaces the missing numbers. 

15. A special type of function calls can appear on the left hand side of the assignment
operator as in > class(x) <- "foo".
```{r}
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
print(R.version.string)
```
