---
title: "RWorksheet_Freires#4b"
output: pdf_document
date: "2024-10-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using Loop Function

for() loop

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value

```{r}
vectorA <- c(1, 2, 3, 4, 5)
zero_matrix <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
  zero_matrix[i, j] <- abs(vectorA[i] - vectorA[j])
  } 
}
print(zero_matrix)

```
2. Print the string "*" using for() function. The output should be the same as shown
in Figure

```{r}
for (i in 1:5) {
  cat(rep('"*"', i), "\n")
}

```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r}
x <- 0
y <- 1

num <- readline(prompt = "Enter the starting number: ")
3
repeat {
  num <- x + y
  if (num > 500) break
  x <- y
  y <- num
  print(num)
}

```

Using Basic Graphics (plot(),barplot(),pie(),hist())

4. Import the dataset as shown in Figure 1 you have created previously.

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result
```{r}
library(readxl)
data_table <- read_excel("/cloud/project/Worksheet#4/data_table.xlsx")
print(head(data_table))
```
b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}
males <- subset(data_table)
females <- subset(data_table)

n_males <- nrow(males)
n_females <- nrow(females)

cat("Number of Male observations: ", n_males, "\n")
cat("Number of Female observations: ", n_females, "\n")
```
