---
title: "RWorksheet_Freires#4a"
output: pdf_document
date: "2024-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. The table below shows the data about shoe size and height. Create a data frame. 

```{r}
shoe_size <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)

height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0)

gender <- c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F",  "M", "F",  "M", "M", "M", "M", "F", "F", "M",  "F", "F", "M", "M",  "F",  "M", "M", "M", "M")

data_table <- data.frame(Shoe_size = shoe_size, Height = height, Gender = gender)
print(data_table)
```

a. Describe the data.

- The data shows the table of Shoe sizes and Height of Male and Female genders. 

b. Create a subset by males and females with their corresponding shoe size and height. What its result? Show the R scripts.

```{r}
males <- subset(data_table, Gender == "M", select = c(Shoe_size, Height))
print(males)
```

```{r}
females <- subset(data_table, Gender = "F", select = c(Shoe_size, Height))
print(females)
```

C. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.

```{r}
mean_shoesize <- mean(shoe_size)
print(mean_shoesize)

```
```{r}
mean_height <- mean(height)
print(mean_height)
```
 
d. Is there a relationship between shoe size and height? Why?
- In my conclusion, there is a relationship because the data has shown that most tall people have bigger shoe sizes. 

2. Construct character vector months to a factor with factor() and assign the result to
factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.
Consider data consisting of the names of months:

"March","April","January","November","January",
"September","October","September","November","August", "January","November","November","February","May","August", "July","December","August","August","September","November","February", April"

```{r}
months_vector <- c("March","April","January","November","January",
"September","October","September","November","August", "January","November","November","February","May","August", "July","December","August","August","September","November","February", "April")
print(months_vector)

factor_months_vector <- factor(months_vector)
print(factor_months_vector)
```

3. Then check the summary() of the months_vector and factor_months_vector. | Interpret the results of both vectors. Are they both equally useful in this case?

```{r}
summary(months_vector)
```
```{r}
summary(factor_months_vector)
```

- My interpretation of the results of months vector is that it shows the length, class, and mode functions. wherein the length is 24, the class is character and the mode is also a character. The results of the factor months vector is showing each month and their levels. Both summary are useful in this case as it has its own function. 

