---
title: "RWorksheet_Freires#4b"
output: pdf_document
date: "2024-10-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using Loop Function

for() loop

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value

```{r}
vectorA <- c(1, 2, 3, 4, 5)
zero_matrix <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
  zero_matrix[i, j] <- abs(vectorA[i] - vectorA[j])
  } 
}
print(zero_matrix)

```
2. Print the string "*" using for() function. The output should be the same as shown
in Figure

```{r}
for (i in 1:5) {
  cat(rep('"*"', i), "\n")
}

```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r}
x <- 0
y <- 1

num <- readline(prompt = "Enter the starting number: ")
3
repeat {
  num <- x + y
  if (num > 500) break
  x <- y
  y <- num
  print(num)
}

```